---
title: "Salmon Counts at Bonneville Dam"
author: "Corinna Hong"
date: "February 18, 2020"
output: html_document
banner: "img/banners/steelhead.jpg"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

<br>

Bonneville Lock and Dam's first powerhouse was completed in 1938 on the Columbia River in Oregon to provide hydropower to the Pacific Northwest (U.S. Army Corps of Engineers, n.d.). This project visualizes changes in steelhead salmon passage counts across Bonneville Dam from 1940 - 2019. It first looks at changes in daily counts, then changes in monthly and yearly counts over the same time period. This project showcases time series data work.


<style>
p.caption {
  font-size: 0.9em;
  font-style: italic;
  color: grey;
  margin-right: 10%;
  margin-left: 15%;  
  text-align: left;
}
</style>


```{r, packages}

# Load packages

library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)

```


```{r}

# Read in data

salmon <- read_csv("cbr_fish_passage_bonneville_allyrs_steelhead.csv") %>% 
  clean_names()

```

```{r}

# Wrangle data for daily counts

salmon_day <- salmon %>%
  unite(year, mm_dd, col = "date", sep = "-") %>% # unite year and mm_dd column and separate values with "-"
  mutate(yr_mo_day = lubridate::parse_date_time(date, "ydm")) %>% # format date
  drop_na(yr_mo_day, value) # drop NAs from value column

```

```{r, fig.align = 'center', fig.cap = "**Daily Steelhead Salmon Passage at Bonneville Dam (1940 - 2019).** Time-seris plot showing daily changes in steelhead salmon counts at Bonneville Dam. Data accessed at: http://www.cbr.washington.edu/dart/query/adult_graph_text"}

# Create graph of daily counts

salmon_day_gg <- ggplot(data = salmon_day, aes(x = yr_mo_day, y = value)) +
  geom_line(color = "forestgreen") +
  labs (x = "Year", y = "Count") +
  scale_y_continuous() +
  theme_minimal()

salmon_day_gg

```

A time-series plot of daily steelhead passage shows changes in the number of steelhead throughout each year. Each year contains a peak where passage counts are the highest. The pattern suggests that passage counts have some seasonality; counts start off low at the beginning of the year, slowly increase until the reach a peak, then being to decrease again.  

```{r}

# Wrangle data for monthly counts

salmon_parse <- salmon_day %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% # pull month from yr_mo_day, label = true gives months 3 letter label
  mutate(year = year(yr_mo_day))

salmon_monthly <- salmon_parse %>% 
  group_by(year, month) %>%
  summarize(total = sum(value)) # Add values column by month for each year

```

```{r, fig.align = 'center', fig.cap = "**Monthly Steelhead Salmon Passage at Bonneville Dam (1940 - 2019).** Time-seris plot showing monthly changes in steelhead salmon counts at Bonneville Dam each year. Counts were totaled by month then plotted for each year. Data accessed at: http://www.cbr.washington.edu/dart/query/adult_graph_text"}

# Graph data for monthly counts

salmon_month_gg <- ggplot(salmon_monthly, aes(x = month, y = total, group = year)) + # separate by year
  geom_line(aes(color = year)) + # color by year
  labs (x = "Month", y = "Count") +
  scale_color_paletteer_c("ggthemes::Blue-Teal") + # change color palette
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) + # Get rid of scientific notation in counts
  theme_minimal()

salmon_month_gg

```

Graphing monthly passage counts at Bonneville Dam zooms in to the pattern of seasonality identified in the daily counts plot. The data show steelhead passage counts are relatively low for most of the year until the summer months. Passage coounts being to increase in the summer months with a peak in August. Separating the data by year suggests an overall trend of an increase in the monthly total counts of steelhead passing though the dam each year.

```{r}

# Wrangle data for yearly counts

salmon_yearly <- salmon_parse %>% 
  group_by(year) %>% 
  summarize(total = sum(value))

```

```{r, fig.align = 'center', fig.cap = "**Annual Steelhead Salmon Passage at Bonneville Dam (1940 - 2019).** Time-seris plot showing annual changes in steelhead salmon counts at Bonneville Dam. Counts were totaled by year. Data accessed at: http://www.cbr.washington.edu/dart/query/adult_graph_text"}

# Graph data for yearly counts

salmon_year_gg <- ggplot(salmon_yearly, aes(x = year, y = total)) +
  geom_line(color = "forestgreen") +
  labs (x = "Year", y = "Count") +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) +
  theme_minimal()

salmon_year_gg


```

Annual steelhead passage counts through Bonneville Dam support the increasing trend seen in the previous plot of monthly counts. This figure shows an overall increasing trend in yearly salmon counts. It also shows dips and spikes in passage counts which could indicate a larger cyclical trend. However, it is hard to fully discern if a cyclical trend exists because the dips and spikes get more pronounced as time passes.

<br>
**References**
<br>
Columbia River DART, Columbia Basin Research, University of Washington. (2019). Adult
Passage Graphics & Text. Available from http://www.cbr.washington.edu/dart/query/adult_graph_text

<br>
Steelhead Trout (NOAA Fisheries). (n.d.). Retrieved from https://www.fisheries.noaa.gov/species/steelhead-trout

<br>
U.S. Army Corps of Engineers. (n.d). *Bonneville Lock & Dam*. Retrieved February 19, 2020, from https://www.nwp.usace.army.mil/bonneville/
