---
title: "Land Use/Land Cover and Watershed Maps"
author: "Corinna Hong"
date: "February 18, 2020"
output: html_document
banner: "img/banners/oahu.jfif"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE,
                      warning = FALSE)
```

testing text length in preview testing text length in preview testing text length in preview testing text length in preview testing text length in preview testing text length in preview testing text length in preview testing text length in preview testing text length in preview testing text length in preview testing text length in preview testing text length in preview testing text length in preview testing text length in preview FIN



```{r}

# Load packages

# General
library(tidyverse)
library(janitor)
library(here)
library(paletteer)

# Spatial
library(sf)
library(tmap)

```

```{r}

# Read in data

land_use <- read_sf(dsn = here("Land_Use_Land_Cover_LULC"), layer = "Land_Use_Land_Cover_LULC")

watersheds <- read_sf(dsn = here("Watersheds"), layer = "Watersheds")

# st_crs(watersheds) checking CRS, both are EPSG: 4326

```

```{r}

# Wrangle data for land cover

land_cover_oahu <- land_use %>%
  dplyr::select(landcover) %>%
  filter(landcover != 0) %>%  # Remove land use category of 0
  st_crop(xmin = -158.31, xmax = -157.62, ymin = 21.26, ymax = 21.75) # Crop to Oahu
  
```

```{r, fig.align = 'center', out.width = '100%'}

# Create map for land use

oahu_map <- tm_basemap("Esri.WorldTopoMap") + # Add basemap
  tm_shape(land_cover_oahu) + # Add land cover
  tm_fill("landcover", title = "Land Cover", alpha = 0.6) # layer elements

tmap_mode("view")

oahu_map


```
<p style="font-size:0.9em; font-style:italic; color: grey">
**Figure 1. Land Use Land Cover Types on Oahu.** Historical land use and land cover classification data as of 1976. Data accessed at: http://geoportal.hawaii.gov/datasets/land-use-land-cover-lulc?geometry=-165.822%2C18.755%2C-149.134%2C22.354
</p>


```{r}

# Wrangle data for watersheds

watersheds_oahu <- watersheds %>% 
  dplyr::select(wuname) %>% 
  st_crop(xmin = -158.31, xmax = -157.62, ymin = 21.26, ymax = 21.75) # Crop to Oahu

```

```{r,  fig.align = 'center', out.width = '100%'}

# Create map of watersheds


oahu_watersheds <- tm_basemap("Esri.WorldTopoMap") +
  tm_shape(watersheds_oahu) +
  tm_fill("wuname", title = "Watershed", alpha = 0.5)


tmap_mode("view")

oahu_watersheds

```
<p style="font-size:0.9em; font-style:italic; color: grey">
**Figure 2. Oahu's Watersheds.** Watersheds of Oahu layered over ESRI World Topomap. Data accessed at: http://geoportal.hawaii.gov/datasets/watersheds?geometry=-165.822%2C18.755%2C-149.134%2C22.354
</p>

<br>
<br>
**References**
<br>
Hawaii Statewide GIS Program. (2017). Land Use Land Cover (LULC). Available from http://geoportal.hawaii.gov/datasets/land-use-land-cover-lulc
<br>
Hawaii Statewide GIS Program. (2017). Surface Water Hydrologic Unit Boundaries (Watersheds). Available from : http://geoportal.hawaii.gov/datasets/watersheds